---
title: "Iconopurrrwafflely"
subtitle: 'Pt I: Making Pretty Basic Waffles'
output: 
  html_notebook: 
    fig_caption: yes
    fig_width: 8
    highlight: zenburn
    number_sections: yes
    theme: cosmo
    toc: yes
    toc_depth: 2
author: Bill Felix
date: October 3, 2017
---

# Setup: plugging in your wafflemaker
```{r, echo=TRUE, message=FALSE, warning=FALSE}
library(tidyverse)                      ## niche R
library(waffle)                         ## wafflemaker
library(extrafont)                      ## icon printer
## at this point, it is important to note that you need to install the FontAwesome font on your machine
## after downloading the zipfile, open and install fontawesome-webfront.tff

loadfonts(device = "win") ## necessary for viewing in RStudio on Windows
```

# Waffles 101

__Pro:__ Really cool right? Often taking the place of pie charts but without the 100% constraint.

__Con:__ Comparing shapes is always tough.

```{r}
## use waffle() from library(waffle)
## plot the ggplot2 object 
(wfle <- waffle(c('Green'=5
                  , 'Orange'=5
                  , 'Blue'=15
                  , 'Pink'=10
                  , 'Lime'=4)
                , title = 'Tasty, Colorful Waffles'
                , xlab = '1 Square = 1 Waffle'
                , rows = 4)
)
```

## Deconstructed Waffles
```{r}
## where named integers form the most basic ingredients
flour <- c('Green'=5)
salt <- c('Orange'=5)
baking_powder <- c('Blue'=15)
sugar <- c('Pink'=10)
vanilla_extract <- c('Lime'=4)

## a vector of named numbers
(mix <- c(flour, salt, baking_powder, sugar, vanilla_extract)
)

## plot
(d_wfle <- waffle(mix
                  , title = 'Tasty, Colorful Waffles'
                  , xlab = '1 Square = 1 Waffle'
                  , rows =4)
)
```

# Plain Ol' Waffles with a Twist  

Adding **glyphs** or **isotypes** to waffle charts allows for an almost infographic feel that can attract an audience. The viewers then can connect the data to objects in real life. Thanks to [waffle](https://github.com/hrbrmstr/waffle/blob/master/R/waffle.R) we have a quick, flexible tool to get these out in the world!

`library(waffle)` is leveraging [Font Awesome](http://fontawesome.io/) for its icons. Loading `library(extrafont)` is necessary to get the icons to appear.  
__Note:__ `loadfonts()` is necessary to view in `Rstudio`!

[Font Awesome Cheatsheet](http://fontawesome.io/cheatsheet/) or `waffle::fa_list()` for help with icons.

## Eggo Minis?
```{r, message=FALSE, warning=FALSE}
## use the built-in argument to reference FontAwesome and its icons
(wfle2 <- waffle(c('Green'=5
                  , 'Orange'=5
                  , 'Blue'=15
                  , 'Pink'=10
                  , 'Lime'=4)
                 , title = 'Tasty, Colorful Eggo Minis'
                 , xlab = '1 Eggo Mini = 1 Waffle'
                 , rows = 4
                 , use_glyph = 'th-large'
                 , glyph_size = 8)
 )
```

## Were they a hit with your kids?
```{r, message=FALSE, warning=FALSE}
(wfle3 <- waffle(c('Green'=5
                  , 'Orange'=5
                  , 'Blue'=15
                  , 'Pink'=10
                  , 'Lime'=4)
                 , rows = 4
                 , use_glyph = 'thumbs-up'
                 , glyph_size = 8)
)
```

## Pictures make everything great right?

Going overboard with icons can surely distract and this example exposes what the consequences of icon overload:

```{r, echo=TRUE, message=FALSE, warning=FALSE}
source('wfles_utils.R')                     ## edit of waffle() elements

### Random Waffles
## copy wfle3 and replace the fontlab vector that is supplied
wfle4 <- wfle3
fa_unicode_vector <- fa_unicode()

## random sample of glyphs
set.seed(77)
n_fontlab <- sample(x = fa_unicode_vector
                    , size = nrow(wfle3$data)
                    , replace =T)

## replace
wfle4$data$fontlab <- n_fontlab
(wfle4)
```
## Crazy useless?
```{r}
wfle5 <- wfle

## replace
set.seed(88)
wfle5$data$value <- sample(wfle5$data$value)
wfle5
```


# What have we made, really?

```{r, message=FALSE, warning=FALSE}
## library(waffle) has iron() for comparisons
iron(
  wfle
  , wfle5
)
```

# How could you make purrrfect waffles?

[If time...](https://cran.r-project.org/web/packages/tidyverse/vignettes/manifesto.html)

Continue on to [Pt II of Iconopurrrwafflely...](advwaffles.nb.html)
